---
title: "CPD Gang Arrests"
author: "Princess Ojiaku"
date: "8/26/2018"
output: html_document
---

```{r setup, warning=FALSE, error=FALSE}

library(readxl)
library(tidyverse)
library(sf)
library(leaflet)
library(viridis)
library(dplyr)


knitr::opts_chunk$set(echo = TRUE)
```

## Read in the CPD Gang Database // police beat shape file

This reads in our data that includes the suspected gang member arrest data, police beat in which the arrest occured, the gang affliation, and the age and race of the suspect. The data goes from 1999 to March of 2018.

Then we read in the shapefile of the CPD police beats, so we can map them. 

```{r read in data, warning=F, error=F}

df_gang_318 <- read_excel("CPD Gang Data/CPD gang database 3-18.xlsx", sheet=1)

# I downloaded the Chicago Police Beats Shapefile from the Chicago data portal
map_filepath <- "Boundaries_Police Beats/geo_export_CPD_BEATS.shp"
cpd_beats <- st_read(map_filepath)

#cleaning the data -- take out any rows with NA for a police beat. 

df_gangs <- filter(df_gang_318, !is.na(O_BEAT))
cpd_beats <- filter(cpd_beats, !is.na(beat_num))

#this is just a quick map to map the police beats 

cpd_beats %>% 
  leaflet() %>% 
  addTiles() %>% 
  addPolygons(popup=~beat_num)
```

#Joining the dataframes and creating a new summary for the map

```{r join data, warning=FALSE, error=FALSE}

# it looks like df_gangs needs a padded 0 in front of some of the numbers to make them 4 digits instead of 3. So let's clean that up before we join it with cpd_beats (which is all 4 digits)

df_gangs$O_BEAT <- str_pad(df_gangs$O_BEAT, 4, pad = "0")

#inner join the two data frames. Used an inner join to drop any rows where the two police beat numbers didn't match up

cpd_beats_gangs<- inner_join(cpd_beats, df_gangs, by=c('beat_num'='O_BEAT'))


#get the number of arrests per police beat and create a new data frame (CPD_gang_map2) with just that info

CPD_gang_map2 <- cpd_beats_gangs %>%
  group_by(beat_num) %>%
  summarize(num_arrests=n()) 



```

## Ploting the CPD data



```{r cpd beats, warning=FALSE, error=FALSE}

#set the color pallette as Blues. 

col_pal <- colorNumeric("Blues", domain=CPD_gang_map2$num_arrests)

#Add the pop-up text with Police Beat and Number of Arrests in that area
popup_sb <- paste0("Police Beat: ", as.character(CPD_gang_map2$beat_num), "<br>Number of arrests: ", as.character(CPD_gang_map2$num_arrests))


# yaassss the following leaflet worked! 

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-87.628598, 41.855372, zoom = 10) %>% 
  addPolygons(data = CPD_gang_map2, 
              fillColor = ~col_pal(CPD_gang_map2$num_arrests), 
              fillOpacity = 0.7, 
              weight = 0.2, 
              smoothFactor = 0.2, 
              popup = ~popup_sb) %>%
  addLegend(pal = col_pal, 
            values = CPD_gang_map2$num_arrests, 
            position = "bottomright", 
            title = "Number of Arrests")

```
